### 2017-11-19 latex/mst_lukkarila_juri_2017.tex
edit

There are some potential error sources and human biases that have likely somewhat affected the results. One factor to keep in mind when analyzing the results is that the people who participated in the testing showed enough interest towards the Conversation Assistant to take part in the testing in the first place, meaning the test participants have to be viewed as inherently positively biased towards the Conversation Assistant in some small degree. It seems probable that someone with no interest towards this type of assistive application would most probably not participate in testing it either. As mentioned, section one and two used different background noise recordings, which might have also affected the results slightly. For the pricing, the price ranges were mistakenly written with inclusive ranges between different options, instead of exclusive ranges, meaning the different options were not completely unambiguous. For example, if a test user wanted to answer \textit{"ten euros"}, there was two possible categories to choose from, which both included this price. It is possible that this could have skewed the pricing results a bit into either direction.

@@

In this work, a conversational assistance application was investigated as a solution to the spoken communication challenges the hearing impaired have. In the investigated method, deaf and hard of hearing individuals are supported in conversational situations by providing real-time speech-to-text transcriptions. A prototype application was implemented for testing the validity and usefulness of the proposed method with current speech recognition technology. A user test was designed and executed for obtaining results to these questions. The user test situation simulated a typical conversation between two persons, but one of the persons had the Conversation Assistant prototype for support, in addition to the hearing augmentation devices they had. Background noise recordings were used to match a realistic acoustic use environment. Realistic background noise was important, since the speech perception ability of an individual with hearing loss typically decreases noticeably in noisy and reverberant acoustic environments. The results indicate that the Conversation Assistant can help hearing impaired people to understand and participate in conversations. \\\\
This first round of user testing also ended up being the only one. At the beginning of the project, two or more rounds of testing were envisioned originally. However, after the first round of testing, more iterations of the same test setup were deemed redundant without significant changes to the Conversation Assistant, which in turn was not possible within the scope of this thesis. Expanding the test setting from a one-on-one conversation to a group conversation would have been the ideal next step, but unfortunately there wasn't enough time for it at this stage. Many test subjects reported that in one-on-one conversations, where they can see the other person well, can be quite easy to follow. Even if they don't hear other person very well, they can use traditional methods like lip-reading. This in turn led many of the test subjects to speculate that the Conversation Assistant would be much more useful in group conversations and in situations where the speaker is farther away or lip-reading is otherwise not possible. Examples given of the latter situations included lectures and video-conferencing meetings. \\\\
During the user testing, one of the test participants had a written language interpreter with her, as she was complete deaf and didn't use any hearing devices. This offered a good opportunity to informally compare the Conversation Assistant side by side with a human, professional speech to text translator. The translator had a laptop computer that was placed next to the Conversation Assistant laptop on the table in front of the test user. The translator had a wireless keyboard that she used to write text to the screen, and the end result was effectively and visually pretty similar to the Conversation Assistant. During the introduction phase and between the test sections, both laptops were simultaneously displaying the translations, the other produced by our ASR system, and the other by the human translator. The human translator paused during the actual testing of course. Based on this brief empirical observation, the Conversation Assistant seemed to be noticeable faster than a human translator, but accuracy was still relatively far off from a human for conversational speech. However, it was interesting to note that the human translation results were far from perfect as well, also containing surprisingly many errors especially in the spelling. The most probable explanation seems that accurate translation has to be sacrificed for speed. All things considered, the Conversation Assistant didn't seem to be too far off from the performance level of a professional human translator. One area, where human translators are by far superior is that they can easily add relevant information to the translation, such as who is speaking (to whom), describing the mood or tone of voice, and also translate other sounds in addition to speech. While arguably a human translator is still better overall in speech to text translation for the hearing impaired, it takes many years of training and practice for a human translator to achieve these results. Also, one person cannot be in many places at the same time and requires monetary compensation for ones time and skills, meaning very few individuals can actually benefit from those translation services in the grand scheme of things. On the contrary, once an excellent ASR model is trained, it can be copied infinitely and used all the time in every place with practically no additional costs, enabling the best achieved level of automatic speech recognition for all very cheaply. \\\\
Ultimately, it is the subjective experience of the end users that matters when evaluating the usefulness of the Conversation Assistant. In fact, typical objective measurements for ASR performance like the word error rate of the speech recognizer might not actually be the most relevant for this specific application: A better word error rate does not automatically mean that the quality of the recognition result is also better in terms of understanding the contents \cite{wang2003word}. Also, the end-goal of the Conversation Assistant is not to have perfect transcriptions of speech, but to support and supplement the user's hearing in real-time conversational situations. If the user can understand everything spoken to them with the help of the system, even though the text is full of errors, then it is succeeding in its intended purpose regardless of the speech recognition accuracy. Consequently, in real-time situations the type of errors is generally much more important than simply the overall percentage, when it comes to understandability and helpfulness \cite{wang2003word}. This observation feels inherently logical for conversations: It is easy to imagine that a few extra letters or words, or a slightly wrong grammatical case might not affect the understandability of a sentence very considerably, whereas a missing key word might render the transcript almost useless. In conversational situations between humans, a lot of information can also be interpreted based on the context and other factors like non-verbal cues \cite{van2005transcription}. This was reported also by our test users, saying that in many cases they could guess the meaning of a sentence well enough from incomplete and erroneous transcripts. Based on these observations, it could be more meaningful to quantify in future work what kind of errors the speech recognizer makes, and more importantly, how each type of error affects the usefulness of the Conversation Assistant. \\\\
On the subject of transcription speed, online ASR is generally far superior to human transcription speed. Human conversion can require more than five times the original audio time, introducing significant latency for the results \cite{gaur2016effects}. Speech recognition very fast \cite{mcgraw2016personalized}. \\\\
Potential objective measurements considered for the user test included eye-tracking, which could have been used to accurately determine the percentage of time the users spend watching the Conversation Assistant screen instead of the person speaking. Eye-tracking could also be used to assess how people tend to use the Conversation Assistant: Do they look back-and-forth quickly between the speaker and the screen, or do they first try to listen a full sentence watching the speaker, and only after that check the screen for help. The User Interfaces research group at Aalto has experience in eye-tracking measurements, and could have provided the necessary equipment if desired. While this information would be interesting from a HCI point-of-view, and have some implications for the UI design and recognition result displaying, it did not answer significantly to the primary and secondary objectives set out for the tests at this stage. It is also possible to obtain a decent approximation for this data by analysing the video recordings made of the test session, which can also show were the test subject was looking at on any given moment during the test. To avoid manually gathering this data from the video, computer vision methods for eye-tracking could be used to get quite accurate results automatically. For example, Krafka et al. used deep convolutional neural networks for successfully tracking eye movement on mobile device cameras \cite{krafka2016eye}. \\\\
Transcripts with error still better than no transcripts at all \cite{shiver2015evaluating}. \\\\
Technology is there but no wide-spread applications yet? \\\\
The application of machine learning techniques has been the driven force behind progress in ASR. The same methods are starting to be applied to other technological solutions for the hearing impaired as well. For example, machine learning techniques have been leveraged to improve the speech signal processing done in hearing aids \cite{goehring2016speech}. \\\\
While not completely solving the spoken communication challenges faced by millions of people, an assistive application like the presented Conversation Assistant prototype could improve the quality of life for hearing impaired persons and the people close to them. In addition, it could contribute towards a more inclusive society and bring about wider socio-economic benefits, such as increased productivity and lower health care expenditure, as a consequence.

@@

More testing. The phrase \textit{"there's no data like more data"} has been often quoted among machine learning practitioners, but it also applies well to user testing. \\\\
Testing the in a group situation with multiple speakers. Staging a group conversation session would bring along many new complications and practical challenges compared to the relatively simple case of having just one test subject and one person running the test. There are also some unresolved technical challenges, like ensuring good quality audio input from all speakers without complicated microphone setups involving multiple microphones. \\\\
One interesting aspect to quantify would be how useful the Conversation Assistant approach is depending on level of hearing loss. Furthermore, whetere deaf and hard of hearing people use the application differently could be investigated as well. This might have some implications for the user interface design, if different sub user groups have different needs and interaction styles. \\\\
Web and mobile application, server-based recognizer. \\\\
Better models for noise-robust conversational speech. \\\\
Microphone methods: access other peoples mobile devices in meetings. Beamforming and directional mics. B-format signal processing for dereverberation and background noise attenuation. \\\\
Speaker diarization. \\\\
Augmented reality text display.

>>>>

There are some potential error sources and human biases that have likely somewhat affected the results. One factor to keep in mind when analyzing the results is that the people who participated in the testing showed enough interest towards the Conversation Assistant to take part in the testing in the first place, meaning the test participants have to be viewed as inherently positively biased towards the Conversation Assistant in some small degree. It seems probable that someone with no interest towards this type of assistive application would most probably not participate in testing it either. As mentioned, section one and two used different background noise recordings, which might have also affected the results slightly. For the pricing, the price ranges were mistakenly written with inclusive ranges between different options, instead of exclusive ranges, meaning the different options were not completely unambiguous. For example, if a test user wanted to answer \textit{"ten euros"}, there was two possible categories to choose from, which both included this price. It is possible that this could have skewed the pricing results a bit into either direction. The test situation used corresponded to one particular type of use scenario, meaning the results obtained may not fully apply to different settings. More testing in different situations, such as in group conversations and lectures would be a logical next step.

@@

In this work, a conversational assistance application was investigated as a solution to the spoken communication challenges the hearing impaired have. In the investigated method, deaf and hard of hearing individuals are supported in conversational situations by providing real-time speech-to-text transcriptions. A prototype application was implemented for testing the validity and usefulness of the proposed method with current speech recognition technology. A user test was designed and executed for obtaining results to these questions. The user test situation simulated a typical conversation between two persons, but one of the persons had the Conversation Assistant prototype for support, in addition to the hearing augmentation devices they had. Background noise recordings were used to match a realistic acoustic use environment. Realistic background noise was important, since the speech perception ability of an individual with hearing loss typically decreases noticeably in noisy and reverberant acoustic environments. The results indicate that the Conversation Assistant can help hearing impaired people to understand and participate in conversations. Based on the results and feedback obtained, the usefulness of the Conversation Assistant seems to depend in large part on the quality and accuracy of the speech recognition results. \\\\
This first round of user testing also ended up being the only one. At the beginning of the project, two or more rounds of testing were originally envisioned. However, after the first round of testing, more iterations of the same test setup were deemed redundant without significant changes to the Conversation Assistant, which in turn was not possible within the scope of this thesis. Expanding the test setting from a one-on-one conversation to a group conversation situation would have been the ideal next step. Many test subjects reported that one-on-one conversations can be relatively easy to follow, when they can see the other person well. This in turn led many of the test subjects to speculate that the Conversation Assistant would be much more useful in group conversations and in situations where the speaker is farther away, or lip-reading is otherwise not possible. Examples given of the latter situations included lectures and video-conferencing meetings. \\\\
During the user testing, one of the test participants brought a written language interpreter, as the person was complete deaf. This offered a good opportunity to informally compare the Conversation Assistant side by side with a human, professional speech to text transcriber. The translator had a laptop computer that was placed next to the Conversation Assistant laptop on the table in front of the test user. The translator had a wireless keyboard that she used to write text to the screen, and the end result was effectively and visually very similar to the Conversation Assistant. During the introduction phase and between the test sections, both laptops were simultaneously displaying transcriptions, the other produced by our ASR system, and the other by the human translator. The human translator paused during the testing. Based on this brief empirical observation, the Conversation Assistant seemed to be noticeable faster than a human translator, but accuracy was still relatively far off from a human translator for conversational speech. However, it was interesting to note that the human translation results were far from perfect as well, containing surprisingly many errors in the spelling. The most probable explanation seems that accurate translation has to be sacrificed for speed. All things considered, the Conversation Assistant didn't appear to be too far off from the performance level of a professional human translator. One area, where human translators are by far superior is that they can easily add relevant information to the translation, such as who is speaking (to whom), describing the mood or tone of voice, and also translate other sounds in addition to speech. While a human translator still seemed to be better overall in speech-to-text translation for the hearing impaired, it takes many years of training and practice for a human translator to achieve these results. Also, one person cannot be in many places at the same time, and requires monetary compensation, meaning only a limited amount of people can benefit from those translation services. On the contrary, once ASR models are trained, they can be copied infinitely and used everywhere with practically no additional costs. This way, utilizing ASR can enable the current best achieved level of automatic speech recognition for all very cheaply. \\\\
Potential objective measurements considered for the user test included eye-tracking, which could have been used to accurately determine the percentage of time the users spend watching the Conversation Assistant screen instead of the person speaking. Eye-tracking could also be used to assess how people tend to use the Conversation Assistant: Do they look back-and-forth quickly between the speaker and the screen, or do they first try to listen a full sentence watching the speaker, and only after that check the screen for help. The User Interfaces research group at Aalto has experience in eye-tracking measurements, and could have provided the necessary equipment if desired. While this information would be interesting from a HCI point-of-view, and have some implications for the UI design and recognition result displaying, it did not answer significantly to the primary and secondary objectives set out for the tests at this stage. It is also possible to obtain a decent approximation for this data by analysing the video recordings made of the test session, which can also show were the test subject was looking at on any given moment during the test. To avoid manually gathering this data from the video, computer vision methods for eye-tracking could be used to get quite accurate results automatically. For example, Krafka et al. used deep convolutional neural networks for successfully tracking eye movement on mobile device cameras \cite{krafka2016eye}. However, this was left for future work due to time constraints. \\\\
Ultimately, it is the subjective experience of the end users that matters when evaluating the usefulness of the Conversation Assistant. Typical objective measurements for ASR performance, like the word error rate may not be the most relevant: A better word error rate does not automatically mean that the quality of the recognition result is also better in terms of understanding the contents \cite{wang2003word}. Also, the end-goal of the Conversation Assistant is not to have perfect transcriptions of speech, but to support and supplement the user's hearing in real-time conversational situations. If the user can understand everything spoken to them with the help of the system, even though the text is full of errors, then it is succeeding in its intended purpose regardless of the speech recognition accuracy. Consequently, in real-time situations the type of errors is generally much more important than simply the overall percentage, when it comes to understandability and helpfulness \cite{wang2003word}. This observation feels inherently logical for conversations: It is easy to imagine that a few extra letters or words, or a slightly wrong grammatical case might not affect the understandability of a sentence very considerably, whereas a missing key word might render the transcript almost useless. In conversational situations between humans, a lot of information can also be interpreted based on the context and other factors like non-verbal cues \cite{van2005transcription}. This was reported also by our test users, saying that in many cases they could guess the meaning of a sentence well enough from incomplete and erroneous transcripts. Based on these observations, it could be more meaningful to quantify what kind of errors the speech recognizer makes, and more importantly, how each type of error affects the usefulness of the Conversation Assistant. \\\\
In recent years, speech recognition has become popular and commonplace among consumer applications. However, very few applications like the one described in this work have been made publicly available so far. Many of the test participants mentioned that they have been interested in this type of assistive application, but there are currently none available that supports Finnish language speech recognition. The technology and software needed for the implementation this type of assistive application is freely available, meaning it would be relatively simple to release one. Previously, one of major challenges for implementing an application like this has likely been acquiring good training data for the speech recognition models, as most transcribed publicly available speech data has been restricted to research purposes, and commercial language resources can be very expensive. Fortunately, companies such as Google now offer application developers a fully working speech recognizer supporting numerous languages. These services generally use server-based speech recognition, like in the case of the \href{https://cloud.google.com/speech/}{Google Cloud Speech API}. \\\\
The application of machine learning techniques has been the driven force behind progress in ASR. The same methods are also starting to be applied to other technological solutions for the hearing impaired. For example, machine learning techniques have been leveraged to improve the speech signal processing done in hearing aids \cite{goehring2016speech} for improved speech perception in noise. While not completely solving the spoken communication challenges faced by millions of hearing impaired people, an assistive application like the presented Conversation Assistant could improve the quality of life for hearing impaired persons, and the people close to them. In addition, it could contribute towards a more inclusive society and bring about wider socio-economic benefits, such as increased productivity and lower health care expenditure, as a consequence.

@@

As mentioned previously, there remains a lot of testing that could be done for further development of the Conversation Assistant. These include objective performance measures for the speech recognition accuracy and speed, which enable the comparison of different speed recognition techniques and implementations. The helpfulness of the Conversation Assistant could also be quantified for the same reasons, for example by counting the number of correctly understood words in the word explanation task. Descriptive evaluation metrics for the understandability and quality of speech recognition results would be very useful in evaluation the Conversation Assistant. One interesting aspect to quantify would be how useful the Conversation Assistant approach is depending on the level of hearing loss. Whether deaf and hard of hearing people use the application differently could be investigated as well. This might have some implications for the user interface design, in the case the are sub user groups with different needs and interaction styles. The phrase \textit{"there's no data like more data"} has been often quoted among machine learning practitioners, but it also applies well to user testing. \\\\ 
For testing with a sensible real-world implementation, a server-based speech recognition system should be implemented with a web interface or mobile application. Comparing on-device and server-based speech recognizers could be done as well. As shown by \cite{mcgraw2016personalized}, it is possible to run an English LVCSR system on smartphone hardware. Investigating the same possibility for Finnish language recognition could be worthwhile in relation to the Conversation Assistant. Using a combination of on-device and server-based recognition could be implemented as well, where a server-based recognizer could be primarily used for better accuracy, but with a locally running recognizer as a back-up when the internet connection is slow or not available. \\\\
Testing the Conversation Assistant in a group situation with multiple speakers was identified as the logical next step. Staging a group conversation session would bring along many new complications and practical challenges compared to the relatively simple case of having just one test user and one person administering the test. There are also some unresolved technical challenges, like ensuring good quality audio input from all speakers without complicated microphone setups involving multiple microphones. Noise-robust conversational speech recognition is arguably critical for the success of the Conversation Assistant. Improving the speech recognition accuracy of the prototype was identified in the user testing to be very important for increased utility of the application. Therefore, new and better speech recognition models for noise-robust conversational speech should be trained for future testing. These models could also take advantage of the latest improvements to the Kaldi toolkit, such as the new \textit{"nnet3"} chain DNN-HMM models, which should be considerable faster for online decoding. Likewise, different microphone configurations and DSP-techniques could be investigated for improved speech recognition accuracy and recognition of different speakers. Speaker diarization, meaning the (visual) separation of recognition results based on the speaker, was mentioned to be desirable by multiple test participants. In group conversation situations, such as meetings, the mobile devices of each participant could be effectively used as wireless microphones. This would work by each person having their device in front of them, and data could be shared through the internet or with Bluetooth. This type of arrangement could enable relatively easy speaker diarization in addition to improved SNR of the speech signals compared to a single microphone. Beamforming and directional microphone arrays could also be used for noise suppression and speaker diarization. Improved techniques for example to directional noise reduction were recently presented in \cite{delikaris2017parametric}. \\\\
The visual dispersion caused by the Conversation Assistant user having to divide their attention between the speaker and the screen displaying text could be reduced with an augmented reality text display. With AR, the text transcriptions could be displayed floating next to speaker for minimal visual dispersion. This could easily be implemented by utilizing for example the \href{https://developer.apple.com/arkit/}{Apple ARKit}.
